Should I Airbnb My Property?

### 2017-03-08 abb_Writeup.Rmd
review - discussion

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of Short Term Preference', fig.lp='xxx'}

@@

Examining the SVM heat maps for the entire metro region under the Actual revenue (left panel, \autoref{potlog}) and the Potential revenue (right panel) show the large increase in possible occupancy rate and nightly rate combinations that lead to short-term preference.  Under a potential revenue scenario, a median occupancy rate and median nightly rate is results in greater than 50\% chance of short-term preference.  Even with very low nightly rates (10th percentile), an occupancy rate around the 65th percentile will achieve short-term preference likelihood in the potential case.

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='\\label{potlog}Logistic Model Results of Short Term Preference', fig.lp='xxx'}

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of Short Term Preferencexxx', fig.lp='xxxx'}

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of Short Term Preference', fig.lp='xxx'}

@@

```{r plot_hostbar, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3, fig.width=4 fig.caption='\\label{htc}Host Type Counts', fig.lp='hostbar'}

@@

Although it gathers considerable interest in both economic and public policy circles, analyses of Airbnb market and their comparison to more traditional long-term leasing strategies remains relatively unaddressed.  This work attempts to fill this void.

The Airbnb market is Melbourne is rapidly expanding, as total listings on the portal have jumped from around 7,000 in 2014 to more than 20,000 in August 2016.  However, of this total, only 6,600 or so of the properties are entire homes or apartments that were listed as available for at least one day during the September 1, 2015 to August 31, 2016 time frame.  Of these, only 4,400 are common bed/bath configurations with non-outlying rates and were booked for at least one night over the period of study. Overall, the Airbnb market is very small relative to the long-term market. We caution future researchers and policy analysts to look deeply at initial estimates of Airbnb counts as many of the listings on the market are either room shares and/or not active.

Comparing annual revenues from the two leasing strategies shows that revenues from more traditional long-term rentals are, on average, higher and much less variable than those from the Airbnb market (Figure 10).  However, there is long right tail to the distibution of short-term revenues suggesting that some owners are doing better, and occasionally much better, on Airbnb than they would on the long-term market. This univariate comparison of revenues does not tell the entire story, as the Airbnb market does not have the same spatial distribution as the long-term market (Figure 8).  Short-term leasing is much more common in the central areas of the city, with small clusters showing up in suburban beach locations and rural tourist destinations like the Yarra Valley wine district.

To control for these differences, we set out to make direct comparison of short and long term revenues on both the short and long term market observations. The shared information that is available on both sets of properties is limited. Our initial work found estimating long term rental values and tenant search times for the short term properties to be feasible with these data constraints while the opposite is much more difficult. In other words, there appear to by many unobserved variables (presense of wi-fi, parking and laundry facilities) and/or none-property specific characteristics (such as minimum stays, pet policies, number of allowed guests, cancellation policies) that influence short term nightly rates and occupancy rates.  As a result, we were not able to accurately estimate likely short-term rates for the long-term observations.

In looking only at the short-term observation with estimated long-term rents we found that approximately one in three Airbnb properties brought in more revenue with its short term strategy than it would have with a long-term approach. As expected, this probability is not equal across space, property type or property size (bed and bath count).  Apartments are much more likely than houses to be more profitable as a short term rental.  Core city properties are more likely than other properties to prefer a short-term strategy, as are larger properties versus smaller ones.  For instance, 35\% of apartments in the city core generated more revenue on Airbnb than they likely would have on the long term market while only 8\% of suburban houses did. Looking at potential revenues -- through eliminating the impact of blocked days -- increases the overall likelihood of short-term preference considerably.

The nightly rates and total occupancies drive the profitability of short-term leasing.  As the nightly rate vs. occupancy rate heat maps and SVM plots (Figures 16, 17, 18 and 19) show, even with very high nightly rates, owners would need to achieve an occupancy rate of at least the 75th percentile in market (around 40\%) in order to achieve a better than 50\% probability of being more profitable in a short-term strategy than a long term one. A 40\% occupancy rate means bookings at least three days out of every week all year long. In Melbourne summers this may be an easy feat as tourist numbers swell, but doing so on a cold week in July likely proves more difficult. The verticalness of the divide between short and long term preference in the SVM plots suggests that occupancy rate is critical to reaching short-term preference in most markets.

The logistic regression models than we estimate show that most of the finding from the univariate and bivariate analyses done earlier are validated (Table 7).  Small apartments in the central city are the best bet for generating greater revenues through short term leasing.  We also test for the impact of host behavior on the probability of short term preference.  We find that more flexibility around the number of guests and the minimum stay increase the probability of short-term preference, while the reverse is true of cancellation policies (more strict equals higher short term preference probability).  As a host, it appears to be best to do what you can to attract bookings and then make it difficult for those bookings to be cancelled.

Testing for the sensitivity of the initial results through consideration of potential revenue figures and by dividing the sample up by likely host motivation, we find the qualitative finding to be generally robust.  While the use of potential revenue does increase the base probability of short-term preference, the same factors -- small, inner city apartment with low minimum stays, strict cancellation and liberal guest capacity -- explain a higher likelihood of short-term preference.

@@

There are a number of limitations to this research.  There are few shared property characterstics between the two datasets, which make cross-estimating rents and rates difficult, especially from the short-term market.  Part of the difficulty in estimating potential short-term rates for long-term properties is that the short term market appears to be heavily influenced by non-property specific factors such as the leasing policies (minimum stay, number of guests, etc.) themselves.  Future research should focus both on building better models to explain (and ultimately predict) short term rates and occupancies as well as to understand the impact of policies such as strict cancellation or no pets on the short term market.  It is likely that this work is being done proprietarily by Airbnb.com itself and other complementary firms such as AirDNA.com, our data provider.

Additionally, we have opted to make comparisons between short and long-term strategies by cross estimating likely rates and rents.  Alternatively, a matching procedure could be employed that would match like properties between the two markets and use comparisons between the two as evidence of potential short and long-term revenues.

@@

>>>>

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Potential Short Term Preference'}

@@

Examining the SVM heat maps for the entire metro region under the Actual revenue (left panel, \autoref{pothm}) and the Potential revenue (right panel) show the large increase in possible occupancy rate and nightly rate combinations that lead to short-term preference.  Under a potential revenue scenario, a median occupancy rate and median nightly rate is results in greater than 50\% chance of short-term preference.  Even with very low nightly rates (10th percentile), an occupancy rate around the 65th percentile will achieve short-term preference likelihood in the potential case.

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE}

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='\\label{pothm}Heatmaps of Potential Short Term Preference'}

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of Potential Short Term Preference'}

@@

```{r plot_hostbar, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3, fig.width=4, fig.caption='\\label{htc}Host Type Counts'}

@@

Although it gathers considerable interest in both economic and public policy circles, analyses of short-term markets and their comparison to more traditional long-term leasing strategies remains relatively unaddressed. The Airbnb market is Melbourne is rapidly expanding as total listings on the portal have jumped from around 7,000 in 2014 to more than 20,000 in August 2016.  However, of this total only 7,400 or so of the properties are entire homes or apartments that were listed as available for at least one day during the September 1, 2015 to August 31, 2016 time frame.  Of these, only 5,788 are common bed/bath configurations with non-outlying rates and were booked for at least one night over the period of study. Overall, the Airbnb market is very small relative to the long-term market. We caution future researchers and policy analysts to look deeply at initial estimates of Airbnb counts as many of the listings on the market are either room shares and/or not active.

Comparing annual revenues from the two leasing strategies shows that revenues from more traditional long-term rentals are, on average, higher and much less variable than those from the Airbnb market (\autoref{revdens}).  However, there is a long right tail to the distibution of short-term revenues suggesting that some owners are doing better, and occasionally much better, in the short-term market than they would on the long-term market. This univariate comparison of revenues does not tell the entire story, as the Airbnb market does not have the same spatial distribution as the long-term market (Figure 8).  Short-term leasing is much more common in the central areas of the city, with small clusters showing up in suburban beach locations and rural tourist destinations like the Yarra Valley wine district.

In analyzing the short-term observations with estimated long-term rents we found that approximately one in four Airbnb properties brought in more revenue with its short-term leasing strategy than it would have with a long-term approach. As expected, this probability is not equal across space, property type or property size (bed and bath count).  Apartments are much more likely than houses to be more profitable as a short-term rental.  Core city properties are more likely than other properties to prefer a short-term strategy, as are smaller properties versus larger ones. For instance, 36\% of apartments in the city core generated more revenue on Airbnb than they likely would have on the long term market while only 8\% of suburban houses did. Looking at potential revenues -- through eliminating the impact of blocked days -- increases the overall likelihood of short-term preference considerably.

The nightly rates and total occupancies drive the profitability of short-term leasing.  As the nightly rate vs. occupancy rate heat maps and SVM plots (\autoref{metrohmp}, \ref{smhm}, \ref{smhmps}, \ref{pothm}) show, even with very high nightly rates, owners would need to achieve an occupancy rate of at least the 75th percentile in market (around 40\%) in order to achieve a better than 50\% probability of being more profitable in a short-term strategy than a long term one. A 40\% occupancy rate means bookings at least three days out of every week all year long. In Melbourne summers this may be an easy feat as tourist numbers swell, but doing so on a cold week in July likely proves more difficult. The vertical-ness of the divide between short- and long-term preference in the SVM plots suggests that occupancy rate is the most critical factor in reaching short-term preference in most markets.

The logistic regression models (Table 4, 6 and 8) than we estimate show that most of the finding from the univariate and bivariate analyses done earlier are validated .  Small apartments in the central city are the best bet for generating greater revenues through short term leasing.  We also test for the impact of host behavior on the probability of short term preference.  We find that more flexibility around the number of guests and the minimum stay increase the probability of short-term preference, while the reverse is true of cancellation policies (more strict equals higher short term preference probability).  As a host, it appears to be best to do what you can to attract bookings and then make it difficult for those bookings to be cancelled.

Testing for the sensitivity of the initial results through consideration of potential revenue figures and by dividing the sample up by likely host motivation, we find the qualitative finding to be generally robust.  While the use of potential revenue does increase the base probability of short-term preference, the same factors -- small inner city apartments with low minimum stays, strict cancellation and liberal guest capacity -- explain a higher likelihood of short-term preference.

@@

There are a number of limitations to this research.  There are few shared property characterstics between the two datasets, which make cross-estimating rents and rates difficult, especially from the short-term market.  Part of the difficulty in estimating potential short-term rates for long-term properties is that the short term market appears to be heavily influenced by non-property specific factors such as the leasing policies (minimum stay, number of guests, etc.) themselves.  Future research should focus both on building better models to explain (and ultimately predict) short-term rates and occupancies as well as to understand the impact of policies such as strict cancellation or no pets on the short term market.  It is likely that this work is being done proprietarily by Airbnb.com itself and other complementary firms such as AirDNA.com, our data provider.

Additionally, we have opted to make comparisons between short- and long-term strategies by imputing long-term rents for short-term properties with a hedonic price model. Alternatively, a matching procedure could be employed that would match like properties between the two markets and use comparisons between the two as evidence of potential short and long-term revenues.

@@

#### Data, Code and Interactivity

Unfortunately both data sources for this work are proprietarily licenses and are not able to be shared by the authors.  All the code used to complete this analysis, done in the R language, is available for download at www.github.com/andykrause/airbnbmelbourne.  An interactive website that allows readers to examine additional facets, breakdown and subsets of the data and the resulting analytical results can be found at shinyapps.io/somename.
